### 2019-10-24 thesis/txt/0-introduction.tex
writing a lot in chapter 2, various figures, data analysis and figures in thesis

\chapter{Introduction}
%Many s These CI builds often produce very long and verbose build logs \mention{add log characteristics, cite what Moritz cited in his proposal?}, stating the progress and results of the various steps within the build.

%These build logs are a highly valuable data source. First of all, for the developers that read them to analyze why their build failed or .. \mention{find more reasons}. Second, for researchers that can harvest the information contained in the logs -- and their metadata -- to study the software engineering process of a project. However they can only utilize the information within the build logs if they can adequately extract the information relevant to them.

%There are many different retrieval techniques use for this task. Beller et al.  used regular expressions when analysing the failure reasons of ruby and Java Maven buildlogs from TravisCI~\cite{beller2017oops}, while Vassallo et al. wrote a custom parser for Java Maven build logs to gather information for build repair hints~\cite{vassallo2018un-break}. Recently Amar et al. greatly reduced the number of build log lines for a developer to inspect by creating a diff between the logs from a failed and successful build~\cite{amar2019mining}. \bp{anecdotal: keyword search?}. Apart from those there are various more retrieval techniques like searching for keywords, ... \todo{more?} \bp{introduce problem: which one to choose}

%With our work we want to support developers, researchers or project managers in deciding which technique is the best one for their use case. 
Continuous Integration (CI) has become a best practice in software engineering.
Many software projects now use CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect bugs more easily~\cite{vasilescu2015quality,duvall2007continuous} and improve developer productivity~\cite{miller2008hundred,hilton2016usage} and communication~\cite{downs2012ambient}.
A build on a CI server typically not only compiles and packages the software, but also executes tests~\cite{beller2017oops} and various kinds of static analyses~\cite{zampetti2017open}.

@@

Currently there is only anecdotal evidence on the performance of these techniques and when a technique should be preferred.
Developers and researches have little support while choosing which technique to use for an retrieval task.
The goal for this thesis is to investigate different information retrieval techniques (IR techniques) and describe under which circumstances we recommend each of the techniques.

We aim to define a model to characterize the different IR techniques as well as the information retrievable from CI build logs.
For our first research question we analyze which criteria influence the suitability of an IR technique.
We support our assumption by implementing and evaluating three techniques:
\begin{itemize}
  \item regular expression synthesis by example using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as TS) and
  \item simple keyword search (referred to as SKWS)
\end{itemize}
Our second research question asks in which cases PBE, TS and SKWS are suited to retrieve information from continuous integration build logs. RQ2 is refined sub-questions along the criteria resulting from RQ1 and compares their instantiations for the three techniques: how many examples the techniques need to perform best, the accuracy of the retrieved output and how diverse the configuring examples can be structurally.
\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which criteria influence the suitability of an information retrieval technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to retrieve information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the retrievals of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}
\end{simplebox}
To evaluate PBE, TS and SKWS we create the \emph{Failing Build Logs Data Set}, which encompasses about 800 log files from 80 repositories. Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural position within the log.

>>>>

\chapter{Introduction}
\plan{finalized, out for review}
Continuous Integration (CI) has become a best practice in software engineering.
Many software projects now use CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect bugs more easily~\cite{vasilescu2015quality,duvall2007continuous} and improve developer productivity~\cite{miller2008hundred,hilton2016usage} and communication~\cite{downs2012ambient}.
A build on a CI server typically not only compiles and packages the software, but also executes tests~\cite{beller2017oops} and various kinds of static analyses~\cite{zampetti2017open}.

@@

Currently there is only anecdotal evidence on the performance of these techniques and when a technique should be preferred.
Developers and researches have little support while choosing which technique to use for a retrieval task.
The goal for this thesis is to investigate different information retrieval techniques (IR techniques) and describe under which circumstances we recommend each of the techniques.

We aim to define a model to characterize different IR techniques as well as the information retrievable from CI build logs.
For our first research question we analyze which criteria influence the suitability of an IR technique.
We support our assumption by implementing and evaluating three techniques:
\begin{itemize}
  \item regular expression program synthesis from examples using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as TS) and
  \item simple keyword search (referred to as SKWS)
\end{itemize}
Our second research question asks in which cases PBE, TS and SKWS are suited to retrieve information from continuous integration build logs. RQ2 is refined into sub-questions along the criteria resulting from RQ1 and compares their instantiations for the three techniques: how many examples a technique needs to perform best, the accuracy of the retrieved output and how diverse the configuring examples can be structurally.
\begin{simplebox}{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] Which criteria influence the suitability of an information retrieval technique for CI build logs?
  \item[\textbf{RQ2:}] When are PBE, TS and SKWS suited to retrieve information from CI build logs?
  \item[\textbf{RQ2.1:}] How many examples do PBE, TS and SKWS need to perform best?
  \item[\textbf{RQ2.2:}] How accurate are the retrievals of PBE, TS and SKWS?
  \item[\textbf{RQ2.3:}] How structurally similar do the examples for PBE and TS need to be for the techniques to be applicable?
\end{itemize}
\end{simplebox}
To evaluate PBE, TS and SKWS we create the \emph{Failing Build Logs Data Set}, which encompasses about 800 log files from 80 repositories. Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural position within the log.

### 2019-10-24 thesis/txt/1-related-work.tex
writing a lot in chapter 2, various figures, data analysis and figures in thesis

\chapter{Related Work}
\label{sec:rw}
\todo{introduction \& glue text}
This section presents various research works adjecent or foundational for our work. Starting with studies about Continuous Integration used in software projects, we move on to past works about build log analysis and their augmentation. We differentiate our work from the more common production runtime log analysis and cover different information retrieval techniques. Lastly this section mentions different Programming by Example resources surrounding Microsoft's work on the PROSE library.

\section{Continuous Integration}
\mention{look into proksch papers}
Various researchers have analyzed industrial and open source logs for failure reasons and their impact on development. Seo et al.~\cite{seo2014programmers} found that few error types such as dependency mismatches are the most prominent cause of build failures at Google. In addition most failures are resolved within two builds. Vassallo et al.~\cite{vassallo2017a-tale} compared open source projects in Java to industrial ones. They determined that testing failures outweigh compilation errors. Open source builds fail most often because of unit tests, whereas release preparations is the primary cause in industrial projects. Beller et al.~\cite{beller2017oops} showed that testing is central to continuous integration when evaluating Travis CI logs for Java and ruby builds. They observed very different kinds, failure rates and numbers of test between programming languages and explained that the low failure rates hint at code being tested before it is sent to the CI server.

@@

\section{Build Log Analysis and Augmentation}
Vassallo et al.~\cite{vassallo2018un-break} tried to shorten the time it takes developers to understand build logs. They summarized relevant information in Maven build logs and augmented them with links to related stack overflow posts. They observed that highlighting the locality and context of an issue is helpful to programmers. We strive to enable a similar summarization by text retrieval while also covering a wider array of programming languages.
\todo{reference their new paper}

@@

\section{Production Log Analysis}
\mention{logpai and drain, [7] K. Fisher and D. Walker. The pads project: an overview. In ICDT,
2011.
[8] K. Fisher, D. Walker, K. Q. Zhu, and P. White. From dirt to shovels:
fully automatic tool generation from ad hoc data. In POPL, 2008.[23] Q. Xi and D. Walker. A context-free markup language for semi- structured text. In PLDI, pages 221–232, 2010.
 The PADS project [7] has en- abled simplification of ad hoc data processing tasks for programmers
by contributing along several dimensions: development of domain specific languages for describing text structure or data format, learn- ing algorithms for automatically inferring such formats [8], and a markup language to allow users to add simple annotations to enable more effective learning of text structure [23] While PADS supports parsing of entire files, FlashExtract allows users to extract only parts of the file thereby avoiding unnecessary complications. PADS’s learner only supports a fixed line-by-line chunking strategy to split the records; in contrast, FlashExtract can learn chunking (aka, struc- ture boundaries) from examples, making it suitable for extracting data fields and records that have arbitrary length (and might cross multiple lines). Finally, PADS primarily targets ad hoc text files. Although one can view webpages and spreadsheet as text files, it is unclear if the PADS learning algorithm can be adapted to work effectively for webpages and spreadsheets.
}

@@

\section{Information Extraction and Retrieval Techniques}
\todo{explain difference here, reference pads and stuff}
\review{Getting the general tot p cic or conceptual information of a text is a common task in information retrieval from semi-structured text sources. Usually this is done by preprocessing the documents, transforming them to a term-by-document matrix. On the matrix we apply a similarity comparison like for example vector space models to calculate the similarity of the different documents to each other~\cite{panichella2016parameterizing}. For our use case, this could be applied by slicing the build logs into lines or small sections. Then similarity measures are used to compare the overall topics in these subparts to the topic of previously labelled logparts.
Instead of receiving a paragraph which is similar to a given query, our works focus more on obtaining specific pieces of texts through regular expressions.}

@@

\section{Program Synthesis by Example}
\todo{how much can/should we explain here? possibly big parts are already explained in related work: vsa, divied and coquer, witness functions, enumeration and ranking (prose theory paper), flash extract specifics like what functions they have (absolute pos, pre/post regex), rankings specific to ci biulg logs?}
\review{
Le et al.~\cite{le2014flashextract:} developed FlashExtract as part of the \emph{Microsoft Program Synthesis using Examples} (PROSE) framework. It is a tool which synthesizes text extraction programs from semi-structured text based on a few given examples. Users can extract multiple fields and structure them with hierarchy and sequence. FlashExtract eliminates the need for the user to understand the entire structure of the processed document and the effort of developing a suitable extraction program themselves.

@@

\section{Semi-Structured Data}
\label{sec:rw-semi-structured-data}
Serge Abiteboul ``characterizes essential aspects of semi-structured data''~\cite{abiteboul1997querying} in the context of integrating data from various data sources.
The content of CI build logs is combined from several tools involved in the build process and several of Abiteboul's aspects can be applied to build logs.

The structure of build logs
\begin{itemize}
  \item is \textbf{implicit}.
  We do not have access to explicit structuring elements or an explicit structure description.
  Computation is required to infer the present structure.
  \item is \textbf{irregular}.
  Changes in the build process or execution environment might change the structure of a build log without notice.
  We observed this in the logs we collected for the \emph{Failing Build Log Data Set}, where for the same repository and build configuration some logs had double new line characters without any noticeable explanation.
  You can see an example of this in Figure \ref{fig:log-4} and Figure \ref{fig:log-5}.
  \item is \textbf{partial}, some parts are highly structured by e.g. structure indicating special characters and others are unstructured, e.g. natural language text in error messages.
  \item can described with an \textbf{a-posteriori data guide} than an a-priori schema.
  There is no fixed specification on how build tools have to structure their output, more we afterwards extract structuring patterns in the produced output.
  \item has a \textbf{rapidly evolving} schema.
  Modifications in the build configuration change the tools involved in the build and therefore the composition of the build log.
\end{itemize}

>>>>

\chapter{Related Work}
\plan{all of RW: 4 - 5 pages left to do + some research!}
\plan{write, 1/3 page, blocked by: other sections}
\label{sec:rw}
\todo{introduction \& glue text}
This section presents various research works adjecent or foundational for our work. Starting with studies about Continuous Integration used in software projects, we move on to past works about build log analysis and their augmentation. We differentiate our work from the more common production runtime log analysis and cover different information retrieval techniques. Lastly this section mentions different Programming by Example resources surrounding Microsoft's work on the PROSE library.

\section{Continuous Integration}
\plan{extend, 1 page}
\mention{look into proksch papers}
Various researchers have analyzed industrial and open source logs for failure reasons and their impact on development. Seo et al.~\cite{seo2014programmers} found that few error types such as dependency mismatches are the most prominent cause of build failures at Google. In addition most failures are resolved within two builds. Vassallo et al.~\cite{vassallo2017a-tale} compared open source projects in Java to industrial ones. They determined that testing failures outweigh compilation errors. Open source builds fail most often because of unit tests, whereas release preparations is the primary cause in industrial projects. Beller et al.~\cite{beller2017oops} showed that testing is central to continuous integration when evaluating Travis CI logs for Java and ruby builds. They observed very different kinds, failure rates and numbers of test between programming languages and explained that the low failure rates hint at code being tested before it is sent to the CI server.

@@

\section{Build Log Analysis and Augmentation}
\label{sec:rw-bl-analysis}
\plan{extend, 1 page}
Vassallo et al.~\cite{vassallo2018un-break} tried to shorten the time it takes developers to understand build logs. They summarized relevant information in Maven build logs and augmented them with links to related stack overflow posts. They observed that highlighting the locality and context of an issue is helpful to programmers. We strive to enable a similar summarization by text retrieval while also covering a wider array of programming languages.
\todo{reference their new paper}

@@

\section{Production Log Analysis}
\plan{research + write, 1/2 - 1 page}
\mention{logpai and drain, [7] K. Fisher and D. Walker. The pads project: an overview. In ICDT,
2011.
[8] K. Fisher, D. Walker, K. Q. Zhu, and P. White. From dirt to shovels:
fully automatic tool generation from ad hoc data. In POPL, 2008.[23] Q. Xi and D. Walker. A context-free markup language for semi- structured text. In PLDI, pages 221–232, 2010.
 The PADS project [7] has en- abled simplification of ad hoc data processing tasks for programmers
by contributing along several dimensions: development of domain specific languages for describing text structure or data format, learn- ing algorithms for automatically inferring such formats [8], and a markup language to allow users to add simple annotations to enable more effective learning of text structure [23] While PADS supports parsing of entire files, FlashExtract allows users to extract only parts of the file thereby avoiding unnecessary complications. PADS’s learner only supports a fixed line-by-line chunking strategy to split the records; in contrast, FlashExtract can learn chunking (aka, struc- ture boundaries) from examples, making it suitable for extracting data fields and records that have arbitrary length (and might cross multiple lines). Finally, PADS primarily targets ad hoc text files. Although one can view webpages and spreadsheet as text files, it is unclear if the PADS learning algorithm can be adapted to work effectively for webpages and spreadsheets.
}

@@

\section{Information Extraction and Retrieval Techniques}
\plan{research + write, 1/2 - 1 page}
\todo{explain difference here, reference pads(maybe that is already in log anlysis!) and stuff}
\review{Getting the general tot p cic or conceptual information of a text is a common task in information retrieval from semi-structured text sources. Usually this is done by preprocessing the documents, transforming them to a term-by-document matrix. On the matrix we apply a similarity comparison like for example vector space models to calculate the similarity of the different documents to each other~\cite{panichella2016parameterizing}. For our use case, this could be applied by slicing the build logs into lines or small sections. Then similarity measures are used to compare the overall topics in these subparts to the topic of previously labelled logparts.
Instead of receiving a paragraph which is similar to a given query, our works focus more on obtaining specific pieces of texts through regular expressions.}

@@

\section{Program Synthesis by Example}
\plan{extend, 1 page}
\todo{how much can/should we explain here? possibly big parts are already explained in related work: vsa, divied and coquer, witness functions, enumeration and ranking (prose theory paper), flash extract specifics like what functions they have (absolute pos, pre/post regex), rankings specific to ci biulg logs?}
\review{
Le et al.~\cite{le2014flashextract:} developed FlashExtract as part of the \emph{Microsoft Program Synthesis using Examples} (PROSE) framework. It is a tool which synthesizes text extraction programs from semi-structured text based on a few given examples. Users can extract multiple fields and structure them with hierarchy and sequence. FlashExtract eliminates the need for the user to understand the entire structure of the processed document and the effort of developing a suitable extraction program themselves.

@@

\todo{maybe scrap this section / incorporate last sentence somewhere else, expalnation will mostly be in rw. How to usefully sturcture the stuff above this?}
In the following we want to give an impression on how the PROSE library and specifically the Flash Extract DSL synthesize programs for extracting substrings from text via regular expressions.
\todo{how much can/should we explain here? possibly big parts are already explained in related work: vsa, divied and coquer, witness functions, enumeration and ranking (prose theory paper), flash extract specifics like what functions they have (absolute pos, pre/post regex), rankings specific to ci biulg logs?}
Our approach is based on the Flash Extract DSL, which in turn is based on the FlashMeta algorithm. Both are described in \todo{ref rw sections}.

\section{Semi-Structured Data}
\plan{finalized, out for review}
\label{sec:rw-semi-structured-data}
Serge Abiteboul ``characterizes essential aspects of semi-structured data''~\cite{abiteboul1997querying} in the context of integrating data from various data sources.
The content of CI build logs is combined from several tools involved in the build process and several of Abiteboul's aspects can be applied to build logs.

The structure of build logs
\begin{itemize}
  \item is \textbf{implicit}.
  We do not have access to explicit structuring elements or an explicit structure description.
  Computation is required to infer the present structure.
  \item is \textbf{irregular}.
  Changes in the build process or execution environment might change the structure of a build log without notice.
  We observed this in the logs we collected for the \emph{Failing Build Log Data Set}, where for the same repository and build configuration some logs had double new line characters without any noticeable explanation.
  You can see an example of this in Figure \ref{fig:log-4} and Figure \ref{fig:log-5}.
  \item is \textbf{partial}. Some parts are highly structured by e.g.\ structure indicating special characters and others are unstructured, e.g.\ natural language text in error messages.
  \item can be described with an \textbf{a-posteriori data guide} rather than an a-priori schema.
  There is no fixed specification on how build tools have to structure their output.
  We afterwards extract structuring patterns in the produced output.
  \item has a \textbf{rapidly evolving} schema.
  Modifications in the build configuration might change the tools involved in the build and therefore the composition of the build log.
\end{itemize}
